---
title: "Lecture02_CodeAlong"
output: html_document
---

## Outline:

-   tidyverse, ggplot, dplyr

-   Demo of plot types

-   Tweaking plots

-   Using PingPongLLM for help

-   dplyr to filter, change, group, and summarize data

## Libraries at the top, always

```{r}
library(tidyverse)
```

## Load data

```{r}
df = read_csv("data/class_survey_simplified.csv")
```

## Get get work

```{r}
df |>
  ggplot(aes(x=comfort_coding)) +
  geom_bar()

```

```{r}
df |>
  filter(language == "Python") |>
  filter(major == "econ") |>
  select(comfort_coding, comfort_math)
```

```{r}
x = 4
x == 4
TRUE + TRUE + FALSE + TRUE
```

```{r}
df |>
  ggplot(aes(x=comfort_coding, fill = language)) +
  geom_bar(position = "dodge")
```

```{r}
df |>
  ggplot(aes(x=comfort_coding, y=comfort_math, color=language)) +
  geom_jitter(width=0.1, height=0.1, size=5, alpha=0.5)
```

```{r}
df |>
  ggplot(aes(x=fake_score, y=language, fill=language)) +
  geom_violin()
```

```{r}
df |>
  ggplot(aes(y=major, fill=major)) +
  labs(
    y = "Major",
    x = "Number of students",
    title = "Stat c131a students by major"
  ) +
  geom_bar()
# ggsave("major_bars.png")
```

## `dplyr` - `group_by` and `summarize`

```{r}
df |>
  group_by(language) |>
  summarize(
    num_students = n(),
    mean_math = mean(comfort_math),
    mean_coding = mean(comfort_coding))
```

You can pipe this straight to ggplot. Note the use of `geom_col` here as we are explicitly setting the y value, while `geom_bar` sets the y to be "the number of rows with this x value"

```{r}
df |>
  group_by(language) |>
  summarize(
    num_students = n(),
    mean_math = mean(comfort_math),
    mean_coding = mean(comfort_coding)) |>
  ggplot(aes(x=language, y=num_students)) +
  geom_col()
```

```{r}
df |>
  summarize(
    n_python = sum(language == "Python"),
    n_r = sum(language == "R"))
```

## Tinkering with a built-in dataset

```{r}
data(Loblolly)
```

```{r}
trees = Loblolly
```

```{r}
trees |>
  ggplot(aes(x=age, y=height, color=Seed)) +
  geom_line()
```

```{r}
trees |>
  group_by(age) |>
  summarize(mean_height = mean(height))
```

### Another built-in dataset

```{r}
ucb = as.data.frame(UCBAdmissions)
```

```{r}
ucb |>
  group_by(Gender, Admit) |>
  summarize(n = sum(Freq))
```

```{r}
ucb |>
  group_by(Gender, Admit) |>
  summarize(n = sum(Freq)) |>
  group_by(Gender) |>
  mutate(fraction = n / sum(n))
```

```{r}
ucb |>
  group_by(Gender, Admit) |>
  summarize(n = sum(Freq)) |>
  group_by(Gender) |>
  mutate(fraction = n / sum(n)) |>
  ggplot(aes(x=Gender, y=fraction, fill=Admit)) +
  geom_col()
```
