---
title: "Lecture03_CodeAlong"
output: pdf_document
author: "YOUR NAME HERE"
---

## Get the data

1.  Go to <https://data.world/wetchler/okcupid-profiles-sf-20120630>
2.  Find the "download" button
3.  Rename the csv to `okcupid.csv`
4.  Upload it to the `data/` folder here ('import' button in Files pane)

## Load libraries

```{r}
library(tidyverse)
```

## Load data

```{r}
okc = read_csv("data/okcupid.csv")
```

## Explore and clean

Sample and slice

```{r}
okc |>
  slice(1:5)
```

```{r}
okc |>
  slice_sample(prop = 0.1)
```

Drop username column

```{r}
okc |>
  select(age, ethnicity)
```

```{r}
okc = okc |>
  select(-username)
```

```{r}
okc = okc |>
  rename(relationship_status = status)
```

Plot, factorize, then replot 'drinks' data

```{r}
okc |>
  group_by(drinks) |>
  summarize(num_people = n(), mean_age = mean(age))
```

```{r}
okc |>
  ggplot(aes(y = drinks)) +
  geom_bar()
```

```{r}
NA
```

```{r}
drinking_order = c(
    "not at all",
    "rarely",
    "socially",
    "often",
    "very often",
    "desperately")
okc = okc |>
  mutate(drinks = factor(drinks, levels=drinking_order))
```

```{r}
okc |>
  drop_na(drinks) |>
  ggplot(aes(y = drinks)) +
  geom_bar()
```

Break down by male/female

```{r}
okc |>
  ggplot(aes(y = sex, fill=drinks)) +
  geom_bar(position = "fill", color="black")
```

Is any age data missing?

```{r}
colSums(is.na(okc))
```

Add boolean "under_30" column, break down drinking this way.

```{r}
okc = okc |>
  mutate(under_30 = age < 30)
```

```{r}
okc |>
  drop_na(drinks) |>
  ggplot(aes(y = under_30, fill=drinks)) +
  geom_bar(position = "fill")
```

Plot income data, then fix -1s to NA

```{r}
okc = okc |>
  mutate(income = if_else(income == -1, NA, income))
```

```{r}
okc |>
  drop_na(income) |>
  group_by(income) |>
  summarize(n=n()) |>
  ggplot(aes(x=income, y=n)) +
  geom_col()
```

Fill body type missing data with "NOT GIVEN"

Plot height data, add vlines for 5 and 6 ft

Change sign column to just keep first word - function - strsplit - sapply

Joining

EDAs - group_by / summarize

-   sign vs diet

-   sign vs drugs

-   age vs drugs

-   Explore on your own - report back interesting finding or two
